
# Importing Necessary Packages:

```{r warning = FALSE}
library(dplyr)
library(readr)
library(tidyr)
library(tidyverse)
library(ggplot2)
library(gapminder)
library(gganimate)
library(gifski)
library(hrbrthemes)
library(patchwork)
library(GGally)
library(viridis)
library(plotly)
# library(MASS)
library(networkD3)
# library(circlize)

```

# EDA: PLAYER STATS

```{r warning = FALSE}

setwd("E:/ARSH/NEU/Spring 2020/DS5110/Projects/Data-Management-and-Processing-Project/datasets/final data")
allplayerdata <- read_csv("cleaned_preprocessed_player_data.csv") %>%
  select(-c(index, PlayerGrade)) 


# allnbadata <- allnbadata %>%
#   mutate(Salary = gsub("\\$", "", Salary),
#          Salary = as.numeric(gsub("\\,", "", Salary)),
#          InflatedSalary = gsub("\\$", "", InflatedSalary),
#          InflatedSalary = as.numeric(gsub("\\,", "", InflatedSalary)),
#          Position = as.factor(Position)) %>%
#   select(-TEAM.y)

```



### EDA w.r.t POSITION:

```{r}

plotCols <- function(data){
  char_cols <- data[,sapply(data, class) == "character"]
  non_char_cols <- data[,sapply(data, class) != "character"]
  
  count = 0
  for(i in char_cols){
    count = count + 1
    g <- ggplot(data, aes(i, fill = Position)) + geom_bar(position = 'dodge') + 
      scale_fill_viridis(discrete = TRUE) +
          theme_ipsum() +
          transition_time(YEAR) +
        ease_aes('linear') +
          theme(
            legend.position="none",
            plot.title = element_text(size=11)
          )  +
          labs(title =  'Year: {frame_time}' ,x = names(char_cols[count]), y = "Count")
      theme(axis.text.x = element_text(angle = 90))
      
      
    anim_save(filename = paste(count,"categorical.gif"), animation = g, path = "animations")
  }
  
 
  return (non_char_cols)
}

```


# Plotting Animated Boxplots:
```{r warning=FALSE}

library(ggplot2)
library(gganimate)
library(lubridate)

non_char_cols <- plotCols(allplayerdata)
non_char_cols <- non_char_cols %>%
  mutate(YEAR = year(as.Date(as.character(allplayerdata$YEAR), format = "%Y")))

colnames_ncc <- colnames(non_char_cols)
count = 0

  for(i in non_char_cols){
      count = count + 1
      g <- allplayerdata %>%
          ggplot(aes(x = as.factor(Position), y = i , fill = Position)) +
          geom_boxplot() +
          scale_fill_viridis(discrete = TRUE) +
          geom_jitter(color="grey", size=0.2, alpha = 0.2 ) +
          theme_ipsum() +
          transition_time(YEAR) +
        ease_aes('linear') +
          theme(
            legend.position="none",
            plot.title = element_text(size=11)
          )  +
          labs(title =  'Year: {frame_time}' ,x = "Position", y = colnames_ncc[count])
          
    anim_save(filename = paste(count,".gif"), animation = g, path = "animations")
  }


```

# Sankey Network Diagram:
```{r}

data_long <- allplayerdata %>%
  group_by(Position) %>%
  select(Position,`OREB`,`REB`,`3PA`,`TOV`,`FTM`,`AST%`) %>%
  gather(key = 'key', value = 'value', -Position) %>%
  sample_n(500) %>%
  ungroup()

colnames(data_long) <- c("source", "target", "value") 
data_long$target <- paste(data_long$target, " ", sep="")

# From these flows we need to create a node data frame: it lists every entities involved in the flow
nodes <- data.frame(name=c(as.character(data_long$source), as.character(data_long$target)) %>% unique())

data_long$IDsource=match(data_long$source, nodes$name)-1 
data_long$IDtarget=match(data_long$target, nodes$name)-1

ColourScal ='d3.scaleOrdinal() .range(["#FDE725FF","#B4DE2CFF","#6DCD59FF","#35B779FF","#1F9E89FF","#26828EFF","#31688EFF","#3E4A89FF","#482878FF","#440154FF"])'

sankeyNetwork(Links = data_long, Nodes = nodes,
              Source = "IDsource", Target = "IDtarget",
              Value = "value", NodeID = "name", 
              sinksRight=FALSE, colourScale=ColourScal, nodeWidth=40, fontSize=13, nodePadding=20)


```


### Types of Rebounds over Time (Stacked)
```{r warning = FALSE}

allplayerdata %>% 
  select(YEAR, REB, DREB, OREB) %>% 
  group_by(YEAR) %>%
  summarise(rebcount = mean(REB),
            offcount = mean(OREB),
            pdefcount = mean(DREB)) %>%
  gather(rebtype, count, -YEAR) %>% 
  ggplot(aes(x = YEAR, y = count, fill = rebtype)) +
  geom_area(aes(ymin = 0, ymax = count, position = "stacked"), alpha = 0.5) +
  scale_fill_discrete(name = "Rebounds",
                      labels = c("Defensive", "Offensive", "Total")) +
  theme_classic() +
  theme(legend.position = "bottom") +
  labs(title = "Types of Rebounds over Time (Stacked)",
       x = "Year",
       y = "Mean")

```


### Types of Points
```{r warning = FALSE}

allplayerdata %>%
  select(YEAR, `3PA`, `3PM`, FTM, FGM, FGA, FGM, FTA) %>%
  group_by(YEAR) %>%
  summarise(tpacount = mean(`3PA`),
            tpmcount = mean(`3PM`),
            ftacount = mean(FTA),
            ftmcount = mean(FTM),
            fgacount = mean(FGA),
            fgmcount = mean(FGM)) %>%
  gather(ptstype, count, -YEAR) %>%
  ggplot(aes(x = YEAR, y = count, fill = ptstype)) +
  geom_ribbon(aes(ymin = 0, ymax = count), alpha = 0.4) +
  scale_fill_discrete(name = "Types of Points",
                      labels = c("FGA", "FGM", "FTA", "FTM",
                                 "TPA", "TPM")) +
  labs(title = "Type of Points",
       x = "Year",
       y = "Count")
       
```

### Three Pointers Made vs Attempted (Stacked)
```{r warning = FALSE}

allplayerdata %>%
  select(YEAR, `3PA`, `3PM`) %>%
  group_by(YEAR) %>%
  summarise(tpacount = mean(`3PA`),
            atpmcount = mean(`3PM`)) %>%
  gather(threepointers, count, -YEAR) %>%
  ggplot(aes(x = YEAR, y = count, fill = threepointers)) +
  geom_area(aes(ymin = 0, ymax = count), alpha = 0.4, position = "stack") +
  scale_fill_discrete(name = "Three Pointers",
                      labels = c("Attempted","Made")) +
  labs(title = "Three Pointers Made vs Attempted (Stacked)",
       x = "Year",
       y = "Mean") +
  theme_classic() +
  theme(legend.position = "bottom")

```

# "Free Throws Made vs Attempted"
```{r warning = FALSE}

allplayerdata %>%
  select(YEAR, `FTM`, `FTA`) %>%
  group_by(YEAR) %>%
  summarise(ftmcount = mean(`FTM`),
            ftacount = mean(`FTA`)) %>%
  gather(freethrows, count, -YEAR) %>%
  ggplot(aes(x = YEAR, y = count, fill = freethrows)) +
  geom_ribbon(aes(ymin = 0, ymax = count), alpha = 0.4) +
  scale_fill_discrete(name = "Free Throws",
                      labels = c("Attempted","Made")) +
  labs(title = "Free Throws Made vs Attempted",
       x = "Year",
       y = "Count")

```

### "Field Goals Made vs Attempted"
```{r warning = FALSE}

allplayerdata %>%
  select(YEAR, `FGM`, `FGA`) %>%
  group_by(YEAR) %>%
  summarise(fgmcount = mean(`FGM`),
            fgacount = mean(`FGA`)) %>%
  gather(fieldgoals, count, -YEAR) %>%
  ggplot(aes(x = YEAR, y = count, fill = fieldgoals)) +
  geom_ribbon(aes(ymin = 0, ymax = count), alpha = 0.4) +
  scale_fill_discrete(name = "Field Goals",
                      labels = c("Attempted","Made")) +
  labs(title = "Field Goals Made vs Attempted",
       x = "Year",
       y = "Count")

```


# Player Stats and Salart over Time:
```{r warning = FALSE}

allplayerdata %>% 
  select(YEAR, MIN, PTS, AST, STL, BLK, TOV, PF, FP, `3PA`, Salary) %>% 
  group_by(YEAR) %>%
  summarise(MIN = mean(MIN),
            PTS = mean(PTS),
            PF = mean(PF),
            TPA = mean(`3PA`),
            Salary = mean(Salary, na.rm = TRUE)) %>%
  gather(pstats, value, -YEAR) %>% 
  ggplot(aes(x = YEAR, y = value, color = pstats)) +
  geom_line(show.legend = FALSE) +
  facet_grid(pstats ~ ., scales = "free") +
  theme_classic() +
  labs(title = "Player Stats over Time",
       x = "Year",
       y = "Mean")

```


# "Game Points over Time (Stacked)"
```{r warning = FALSE}

allplayerdata %>%
  select(YEAR, `FP`, `3PM`, PTS, FGM, FTM) %>%
  group_by(YEAR) %>%
  summarise(zptscount = mean(PTS),
            fgmcount = mean(`FGM`),
            atpmcount = mean(`3PM`),
            bftmcount = mean(FTM)) %>%
  gather(points, count, -YEAR) %>%
  ggplot(aes(x = YEAR, y = count, fill = points, position = "stack")) +
  geom_area(aes(ymin = 0, ymax = count), alpha = 0.4) +
  scale_fill_discrete(name = "Points",
                      labels = c("Field Goals Made",
                                 "Free Throws Made",
                                 "Three Pointers Made",
                                 "Points Scored")) +
  labs(title = "Game Points over Time (Stacked)",
       x = "Year",
       y = "Mean") +
  theme_classic() +
  theme(legend.position = "bottom")

```



### Types of Rebounds over Time (Stacked)
```{r warning = FALSE}

allplayerdata %>% 
  select(YEAR, REB, DREB, OREB) %>% 
  group_by(YEAR) %>%
  summarise(rebcount = mean(REB),
            offcount = mean(OREB),
            pdefcount = mean(DREB)) %>%
  gather(rebtype, count, -YEAR) %>% 
  ggplot(aes(x = YEAR, y = count, fill = rebtype)) +
  geom_area(aes(ymin = 0, ymax = count, position = "stacked"), alpha = 0.5) +
  scale_fill_discrete(name = "Rebounds",
                      labels = c("Defensive", "Offensive", "Total")) +
  theme_classic() +
  theme(legend.position = "bottom") +
  labs(title = "Types of Rebounds over Time (Stacked)",
       x = "Year",
       y = "Mean")

```


### Types of Points
```{r warning = FALSE}

allplayerdata %>%
  select(YEAR, `3PA`, `3PM`, FTM, FGM, FGA, FGM, FTA) %>%
  group_by(YEAR) %>%
  summarise(tpacount = mean(`3PA`),
            tpmcount = mean(`3PM`),
            ftacount = mean(FTA),
            ftmcount = mean(FTM),
            fgacount = mean(FGA),
            fgmcount = mean(FGM)) %>%
  gather(ptstype, count, -YEAR) %>%
  ggplot(aes(x = YEAR, y = count, fill = ptstype)) +
  geom_ribbon(aes(ymin = 0, ymax = count), alpha = 0.4) +
  scale_fill_discrete(name = "Types of Points",
                      labels = c("FGA", "FGM", "FTA", "FTM",
                                 "TPA", "TPM")) +
  labs(title = "Type of Points",
       x = "Year",
       y = "Count")
       
```


### Three Pointers Made vs Attempted (Stacked)
```{r warning = FALSE}

allplayerdata %>%
  select(YEAR, `3PA`, `3PM`) %>%
  group_by(YEAR) %>%
  summarise(tpacount = mean(`3PA`),
            atpmcount = mean(`3PM`)) %>%
  gather(threepointers, count, -YEAR) %>%
  ggplot(aes(x = YEAR, y = count, fill = threepointers)) +
  geom_area(aes(ymin = 0, ymax = count), alpha = 0.4, position = "stack") +
  scale_fill_discrete(name = "Three Pointers",
                      labels = c("Attempted","Made")) +
  labs(title = "Three Pointers Made vs Attempted (Stacked)",
       x = "Year",
       y = "Mean") +
  theme_classic() +
  theme(legend.position = "bottom")

```

# "Free Throws Made vs Attempted"
```{r warning = FALSE}

allplayerdata %>%
  select(YEAR, `FTM`, `FTA`) %>%
  group_by(YEAR) %>%
  summarise(ftmcount = mean(`FTM`),
            ftacount = mean(`FTA`)) %>%
  gather(freethrows, count, -YEAR) %>%
  ggplot(aes(x = YEAR, y = count, fill = freethrows)) +
  geom_ribbon(aes(ymin = 0, ymax = count), alpha = 0.4) +
  scale_fill_discrete(name = "Free Throws",
                      labels = c("Attempted","Made")) +
  labs(title = "Free Throws Made vs Attempted",
       x = "Year",
       y = "Count")

```

### "Field Goals Made vs Attempted"
```{r warning = FALSE}

allplayerdata %>%
  select(YEAR, `FGM`, `FGA`) %>%
  group_by(YEAR) %>%
  summarise(fgmcount = mean(`FGM`),
            fgacount = mean(`FGA`)) %>%
  gather(fieldgoals, count, -YEAR) %>%
  ggplot(aes(x = YEAR, y = count, fill = fieldgoals)) +
  geom_ribbon(aes(ymin = 0, ymax = count), alpha = 0.4) +
  scale_fill_discrete(name = "Field Goals",
                      labels = c("Attempted","Made")) +
  labs(title = "Field Goals Made vs Attempted",
       x = "Year",
       y = "Count")

```


# Player Stats and Salart over Time:
```{r warning = FALSE}

allplayerdata %>% 
  select(YEAR, MIN, PTS, AST, STL, BLK, TOV, PF, FP, `3PA`, Salary) %>% 
  group_by(YEAR) %>%
  summarise(MIN = mean(MIN),
            PTS = mean(PTS),
            PF = mean(PF),
            TPA = mean(`3PA`),
            Salary = mean(Salary, na.rm = TRUE)) %>%
  gather(pstats, value, -YEAR) %>% 
  ggplot(aes(x = YEAR, y = value, color = pstats)) +
  geom_line(show.legend = FALSE) +
  facet_grid(pstats ~ ., scales = "free") +
  theme_classic() +
  labs(title = "Player Stats over Time",
       x = "Year",
       y = "Mean")

```


# "Game Points over Time (Stacked)"
```{r warning = FALSE}

allplayerdata %>%
  select(YEAR, `FP`, `3PM`, PTS, FGM, FTM) %>%
  group_by(YEAR) %>%
  summarise(zptscount = mean(PTS),
            fgmcount = mean(`FGM`),
            atpmcount = mean(`3PM`),
            bftmcount = mean(FTM)) %>%
  gather(points, count, -YEAR) %>%
  ggplot(aes(x = YEAR, y = count, fill = points, position = "stack")) +
  geom_area(aes(ymin = 0, ymax = count), alpha = 0.4) +
  scale_fill_discrete(name = "Points",
                      labels = c("Field Goals Made",
                                 "Free Throws Made",
                                 "Three Pointers Made",
                                 "Points Scored")) +
  labs(title = "Game Points over Time (Stacked)",
       x = "Year",
       y = "Mean") +
  theme_classic() +
  theme(legend.position = "bottom")

```

```{r}

setwd("E:/ARSH/GitHub/Data-Management-and-Processing-Project/datasets/final data")

player_data <- read_csv("cleaned_nba_player_data.csv")

nba_data <- read_csv("cleaned_nba_player_team_data.csv")

team_stats <- read_csv("team_stats_viz.csv")

```

```{r warning = FALSE}

target2 <- c("WAS", "DAL", "UTA", "POR", "TOR", "MIA") # teams with top 3P% 


# Salary Over the years 
team_salary_yrs <- nba_data %>%
  filter(TEAM %in% target2) %>%
  group_by(YEAR, TEAM) %>%
  summarise(total_salary = sum(Salary, na.rm = T), win_perct = mean(`team_WIN%`))


team_salary_yrs_plot <- ggplot(data = team_salary_yrs, mapping = aes(x=YEAR,y=total_salary)) + 
  geom_point(aes(color = TEAM, size = win_perct, alpha =1/10)) + geom_smooth() + 
  theme(title = element_text(color = "Black"), plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  ggtitle("Teams Salary Improvement") +
  xlab("YEAR") + ylab("Team Salary") 
team_salary_yrs_plot

```

```{r warning = FALSE}


###### 3PM over years Vs Salary
team_win_yrs <- nba_data %>%
  filter(TEAM %in% target2) %>%
  group_by(YEAR, TEAM) %>%
  summarise(win_perct = mean(`team_WIN%`), total_3pm = sum(team_3PM, na.rm = T))

team_win_salary_yrs_plot <- ggplot(data = team_win_yrs, mapping = aes(x=YEAR, y=total_3pm)) + 
  geom_point(aes(color = TEAM, size = win_perct, alpha =1/10)) + geom_smooth() + 
  theme(title = element_text(color = "Black"), plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  ggtitle("3 Pointer Along the Years") +
  xlab("YEAR") + ylab("3 Pointer Field Goals Made")
team_win_salary_yrs_plot

```
# TEAMS SALARY VS. OTHER FEATURES
```{r warning = FALSE}

top_bottom_teams_salary_3PA <- ggplot(data = team_stats, mapping = aes(x=team_3PA,y=Salary)) + 
  geom_point(aes(color = TEAM), show.legend = F)  + geom_smooth()+ ylab("Salary") + xlab("3PA") + ggtitle("Teams 3 point Attempt Vs. Salary") + theme(legend.position = "bottom")
top_bottom_teams_salary_3PA


top_bottom_teams_salary_FGM <- ggplot(data = team_stats, mapping = aes(x=team_FGM, y = Salary)) + geom_point(aes(color = TEAM), show.legend = F) + geom_smooth()+ ylab("Salary") + ggtitle("Teams Field Goals Made Vs. Salary") + xlab("FGM")
top_bottom_teams_salary_FGM


top_bottom_teams_salary_PTS <- ggplot(data = team_stats, mapping = aes(x=team_PTS, y = Salary)) + geom_point(aes(color = TEAM), show.legend = F) + geom_smooth() + ylab("Salary") + xlab("Teams Points") + ggtitle("Teams Points Vs. Salary")
top_bottom_teams_salary_PTS


top_bottom_teams_salary_OREB <- ggplot(data = team_stats, mapping = aes(x=team_OREB, y = Salary)) + geom_point(aes(color = TEAM), show.legend = F) + geom_smooth() + ylab("Salary") + xlab("Teams OREB") + ggtitle("Teams Offensive Rebounds Vs. Salary")
top_bottom_teams_salary_OREB


top_bottom_teams_salary_DREB <- ggplot(data = team_stats, mapping = aes(x=team_DREB, y = Salary)) + geom_point(aes(color = TEAM), show.legend = F) + geom_smooth() + ylab("Salary") + xlab("Teams DREB") + ggtitle("Teams Defensive Rebounds Vs. Salary")
top_bottom_teams_salary_DREB


top_bottom_teams_salary_PF <- ggplot(data = team_stats, mapping = aes(x=team_PF, y = Salary)) + geom_point(aes(color = TEAM), show.legend = F) + geom_smooth()+ ylab("Salary") + xlab("Teams Fouls") + ggtitle("Teams Fouls Vs. Salary")
top_bottom_teams_salary_PF

```

